---
title: "Supplementary Analysis and Visualization"
author: ["Shixiang Wang", "Huimin Li", "Xuan Wang", "Minfang Song", "Xue-Song Liu (Corresponding author)"]
date: "`r Sys.Date()`"
output:
  rmdformats::readthedown:
    highlight: kate
    self_contained: false
    mathjax: true
    lightbox: true
    gallery: true
bibliography: ref.bib
link-citations: yes
---

```{r knitr_init_99, echo=FALSE}
library(knitr)
library(rmdformats)
## Global options
options(max.print = "75")
opts_chunk$set(
  echo = TRUE,
  comment = "#>",
  collapse = TRUE,
  cache = TRUE,
  dpi = 300,
  warning = FALSE,
  message = FALSE,
  tidy = "styler"
)
opts_knit$set(width = 75)
```

This part shows some supplementary analyses and descriptions.

## FACETS results

We used 2 methods (Sequenza and FACETS) for copy number calling.
However, only Sequenza was used for previous downstream analyses due to
the fact that FACETS generated some unreasonable copy number profile for a few samples.
We will talk it here.

Let's load the packages and signature data from FACETS and Sequenza.

```{r}
library(sigminer)
library(ggplot2)

load("../output/Sig.CNV.facets.W.RData")
load("../output/Sig.CNV.seqz.W.RData")
load("../output/EST.facets.W.all.RData")
```

According to the estimation of signature number, 6 signatures were extracted.

```{r}
p <- show_sig_number_survey(EST.facets.W.all, right_y = NULL)
add_h_arrow(p, x = 6.1, y = 0.986, seg_len = 0.5)
```

The result signatures are shown as the below.

```{r fig.width=12, fig.height=7}
show_sig_profile(Sig.CNV.facets.W, style = "cosmic", method = "W", normalize = "feature") +
  labs(caption = "Signatures from FACETS")
```

FACETS result has one special and extra signature (**Sig1**) when compared to [Sequenza result](#copy-number-signature-profile).
The **Sig1** shows copy number segments with high percentage of component "copy number value = 0".

> Other signatures are not identical but very similar in most features beteween FACETS and Sequenza results.


We check some samples with **Sig1** enriched.


```{r, message=TRUE}
cn_groups = get_groups(Sig.CNV.facets.W, method = "consensus", match_consensus = TRUE)
cn_expo = get_sig_exposure(Sig.CNV.facets.W, type = "relative")
# Focus on Sig1
df = dplyr::left_join(cn_groups, cn_expo)
load("../output/CNV.facets.RData")
```

```{r, message=TRUE, fig.height=6, fig.width=14}
samps_to_show <- df %>%
  dplyr::filter(enrich_sig == "Sig1") %>%
  dplyr::arrange(dplyr::desc(Sig1)) %>%
  dplyr::slice(1:6) %>%
  dplyr::pull(sample)

samps_to_show

show_cn_profile(
  data = CNV.facets, chrs = paste0("chr", c(1:22, "X")), nrow = 3, ncol = 2, show_title = T,
  samples = samps_to_show
)
```

We can observe that the samples with **Sig1** enriched have many homozygous deletions, this is unreasonable.
I made a communication with the author of FACETS ([issue link](https://github.com/mskcc/facets/issues/147)),

>  So only logical explanation is homozygous deletion. However they are too many of them and so I believe this is assay artifact rather than real.

This problem seems to come from the raw data. However, we checked the data and this signature cannot be found
from the result of Sequenza.

To summary, we used result from Sequenza instead of FACETS for downstream analyses because we found there 
may be some unknown method problems in FACETS (in general, this method is still reliable).

## Method comparison between Macintyre et al and our study

In section ["Tally variation records"](#tally-variation-records), I said that our method (defined as "Wang" or "W")
has some advantages over the method used in Macintyre et al (defined as "Macintyre" or "M").

I will talk the details here.

Firsty let's load the results from "M" method (the code to generate the data can be found [here](https://github.com/ShixiangWang/prad_signature/blob/master/analysis/src/02-identify-copynumber-signatures.R)).

```{r}
load("../output/Sig.CNV.seqz.M.RData")
load("../output/EST.seqz.M.RData")
load("../output/CNV.seqz.tally.M.RData")
```

### Sample-by-component matrix generation

The key difference between method "W" and method "M" is the way to generate the sample-by-component matrix.

- Method "W" defines the components and counts them (details see section ["Tally variation records"](#tally-variation-records)).
- Method "M" applies mixture modeling to separate the 6 copy number features distributions into mixtures of Poisson or Gaussian distributions. 

Here we found that method "M" has the following drawbacks:

1. the computation needs huge computing resources. For generating the matrix used for NMF algorithm, it
takes about 5100s (~1.5h) with 20 cores for method "M" while only 86s (~1.5m) with 20 cores for method "W".
2. the number of component is dependent on the raw data. If we change the sample size or apply the method "M"
to another cohort, the number of component may change. Method "W" will not be affected by this because it uses a
fixed number of component.
3. the distribution parameters for each distribution are not identical for different runs.
Even for the same data, the result distributions and sample-by-component matrix may change.

At the start of this study, I used the method "M" for all analyses. 
Finally, we created method "W" based on the
use and rethinking of method "M" and SBS signature extraction method from previous reports.

Next section, I will talk more problems raised by method "M" from the visualization view.

### Copy number signatures from "M" method


We still use the cophenetic plot for deciding on the signature number.

```{r}
show_sig_number_survey(EST.seqz.M, right_y = NULL)
```

We selected 4 signatures. Let's see their signature profile.


```{r}
show_sig_profile(Sig.CNV.seqz.M, normalize = "column", style = "cosmic", 
                 paint_axis_text = FALSE)
```

This is a bar plot similar to the one reported by Macintyre et al.

This plot raises some issues.

#### Issue 1

The matrix used to generate the plot is column normalized, so **the heights of bar cannot be compared with each other**.

For example, firstly we see the **Sig1**, we may conclude that **component "bp10MB2" has more contributions than component "bp10MB1"**. But if we see the raw matrix, we can see that this is wrong!

We cannot compare two or more componets in a row when we normalize the data by column! The plot above may misguide the readers.
For the plot above, we can only compare the same component across all signatures. For example,
we can conclude that **Sig4** captures the majority of component **"segsize12"** and **"segsize11"**. 


```{r}
show_sig_profile(Sig.CNV.seqz.M, normalize = "raw", style = "cosmic", 
                 paint_axis_text = FALSE)
```

To avoid misguiding readers, gradient colors may work. Also, we must tell readers in text that
we should compare one component in a column insteads two or more componets in a row.

```{r}
show_sig_profile(Sig.CNV.seqz.M, normalize = "column", style = "cosmic", 
                 paint_axis_text = FALSE, set_gradient_color = TRUE)
```

#### Issue 2

From plot above, we can only see different components, but we cannot see the meaning of each component.
For method "M", each component represents a distribution. 

For example, what's the meaning of the component "copynumber2"? To fix this issue, we tried adding
the key parameter value in the plot.


```{r}
show_sig_profile(Sig.CNV.seqz.M, normalize = "column", style = "cosmic", 
                 paint_axis_text = FALSE, set_gradient_color = TRUE, 
                 params = CNV.seqz.tally.M$parameters, y_expand = 1.5)
```

According to the text just added, we can see that component "copynumber2" represents a Gaussian
distribution with mean is 2; component "copynumber1" represents a Gaussian
distribution with mean is 0.84. 

#### Issue 3

Even with the plot above, we cannot know the dispersion of each component.
All distributions with weight for each copy number feature can be viewed as the below.

```{r}
show_cn_components(CNV.seqz.tally.M$parameters, show_weights = TRUE)
```

### Why we created method "W"

Here, I want to talk something more.

If we take case of all available values for each features, we would notice that there are only
limited choices. Let's visualize them.

```{r}
load("../output/CNV.seqz.tally.W.RData")
```

```{r, fig.width=6, fig.height=4}
show_cn_features(CNV.seqz.tally.W$features[1:6], method = "W", log_y = TRUE)
```

Naturally we can divide them into categories and get a fixed number of components.
It would be easy to read, understand and compare.
Look back this part, why we choose a complicated method with many drawbacks instead of a simple one?

