---
title: "Prostate Cancer Variation Signature Analysis Report"
author: ["Shixiang Wang", "Huimin Li", "Minfang Song", "Xuan Wang", "Xue-Song Liu (Corresponding author)"]
date: "`r Sys.Date()`"
output:
  rmdformats::readthedown:
    highlight: kate
    self_contained: true
    mathjax: true
#    thumbnails: true
    lightbox: true
    gallery: true
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding, output_dir = "../docs") })
bibliography: ref.bib
link-citations: yes
---

Contact Prof. Xue-Song Liu: <liuxs@shanghaitech.edu.cn>

This report is written to help user understand what and how we did in this project. 
Please read the formal manuscript **TBD** for more details.

> This document is compiled from an Rmarkdown file which contains all code or description necessary to (auto-)reproduce the analysis for the accompanying project. Each section below describes a different component of the analysis and all numbers and figures are generated directly from the underlying data on compilation.

# Variation Summary

```{r, child="01-variation-summary.Rmd"}

```

# Signature Profile

```{r, child="02-signature-profile.Rmd"}

```

# Association Analysis

```{r, child="03-association-analysis.Rmd"}

```

# Survival Analysis

```{r, child="04-survival-analysis.Rmd"}

```


# Supplementary Analysis and Visualization

```{r, child="99-supp.Rmd"}

```


# Reference

